---
title: "Pairwise comparisons analysis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
params: 
  test: "t.test"
  pval: 0.05
  outliers: FALSE
  data_table: NULL
  labels: NULL
  grouping: NULL
 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Analysis summary

The significance threshold for pairwise comparisons was set to p-val < ``r params$pval``. \
Only comparisons with a p-value below this cut-off are included in the results file. \
All p-values are two-tailed and no adjustment for multiple comparisons was performed.\

```{r, echo = FALSE, results='asis'}
if(params$test == "t.test"){
    
    #Pval for variane ratio test
    pvar_test = data.frame(variable = paste(params$labels[,1], params$labels[,2]),
                           pval_var = rep(NA, ncol(params$data_table)), 
                           test = rep(NA, ncol(params$data_table)))
    
    
    for(i in 1:ncol(params$data_table)){
      p.var = stats::var.test(params$data_table[,i]~params$grouping)$p.value
      
      pvar_test$pval_var[i] = p.var  
    }    
    
    pvar_test$test = ifelse(pvar_test$pval_var<0.05, "Welch's t-test", "T-test")
    
    cat("You selected a t-test. Therefore, beeRapp automatically checked it the variance
        between groups is equal using the var.test function. If yes, an unpaired t-test was performed. If variance was unequal, Welch's t-test was performed, instead. The table below shows which type of test was performed for each variable. \n")
   
    
  datatable(pvar_test[,c(1,3)], extensions = c('FixedColumns', "Scroller", "Buttons"), options = list(
  dom = 'Bfrtip',
  deferRender = TRUE,
  scrollY = 200,
  scroller = TRUE,
  buttons = "excel",
  scrollX = TRUE
))
    
}

if(params$test == "wilcox"){
    cat("Pairwise comparisons were performed with a non-parametric Wilcoxon rank sum test.")
}
if(params$test == "permutation-test"){
    cat("Pairwise comparisons were performed with a permutation based test as implemented in the independence_test function from the coin R package.")
}
```

\

```{r, echo = FALSE, results='asis'}

if(params$outliers == TRUE){
    cat('###', "Outlier analysis", '\n')
    cat("You chose to perform and outlier screening and remove outliers from the analysis.To this end, beeRapp first transformed each measurement to a z-score by subtracting the group mean and dividing by the standard deviation for each variable. Z-scores exceeding 3.29 in absolute value (p<0.001) were considered outliers and removed from the analysis.", '  \n', '  \n')
    cat('#####', 'Outlier detection results', '  \n')
    
    
    #Create a data table for outliers
    outliers_df = data.frame(variable = paste(params$labels[,1], params$labels[,2]),
                             outliers =  rep(NA, ncol(params$data_table)))
    
    
    #Transform grouping to a factor
    grouping = factor(params$grouping)
    
    for (i in 1:ncol(params$data_table)){
      
      y_var = params$data_table[,i]
      names(y_var) = rownames(params$data_table)
      group1 = y_var[grouping == base::levels(grouping)[1]]
      group2 = y_var[grouping == base::levels(grouping)[2]]
      
      #Transform values to z-scores
      group1_z = (group1 -  base::mean(stats::na.omit(group1)))/stats::sd(stats::na.omit(group1))
      group2_z = (group2 - base::mean(stats::na.omit(group2)))/stats::sd(stats::na.omit(group2))
      
      #get potential outliers
    outliers1 = base::names(na.omit(group1_z))[base::abs(na.omit(group1_z))>3.29]
    outliers2 = base::names(na.omit(group2_z))[base::abs(na.omit(group2_z))>3.29]
    outliers = c(outliers1, outliers2)
      
    
      if(length(outliers > 0)){
        
        outliers_df$outliers[i] = paste(outliers, collapse = ",")
    }
      
    
    }

          if(all(is.na(outliers_df$outliers))){
              cat("There were no outliers detected in any of the variables.")
          
          }else{
          cat("The table below provides an overview of detected outliers. The values
              correspond to rownames in the original data table that were flagged as outliers and removed from the analysis for each variable.", "\n")
              
        datatable(outliers_df, extensions = c('FixedColumns', "Scroller", "Buttons"), options = list(
  dom = 'Bfrtip',
  deferRender = TRUE,
  scrollY = 200,
  scroller = TRUE,
  buttons = "excel",
  scrollX = TRUE
))
      }
 
}       
```


### Session info
This analysis was performed with beeRapp's 'Boxplots' module. \
The following R version and packages were used for the analysis.

```{r, echo = FALSE}
if(params$test == "permutation-test"){
    info = c(paste ("R", paste0(R.Version()[c("major","minor")], collapse = ".")),
    paste("ggplot2", packageVersion("ggplot2")),
    paste("stats", packageVersion("stats")),
    paste("coin", packageVersion("coin")))
    print(info)
}else{
    info = c(paste ("R", paste0(R.Version()[c("major","minor")], collapse = ".")),
    paste("ggplot2", packageVersion("ggplot2")),
    paste("stats", packageVersion("stats")))
    print(info)
}

```



